---
title: "Maps or Geo-Spatial Data Visualization"
output: rmarkdown::github_document
---

https://www.youtube.com/watch?v=AbrReroOWS8

```{r}
if (!require("pacman")) install.packages("pacman")
library("pacman")


knitr::opts_chunk$set(
  fig.path = "figures/tolle-karte/"
)
```

## Map Data

The ```maps``` package provides the coordinates of the borders for each country in the world.

```{r}
p_load(ggplot2, maps)
germany <- map_data("world", region = "Germany")

p_load(knitr)
kable(germany[1:5,])
```
```coord_map``` of the ```mapproj``` package, allows geographical projection of the data.

```{r germany-map}
p_load(mapproj)
p_load(dplyr)

germany %>% ggplot(aes(long, lat, group = group, fill = subregion)) +
  geom_polygon(color = "black") +
  coord_map()
```

```{r germany-map-albers}
germany %>% ggplot(aes(long, lat, group = group, fill = subregion)) +
  geom_polygon(color = "black") +
  coord_map("albers",  lat0 = 45.5, lat1 = 29.5)
```

## Spatial Data with sf

Some countries provide spatial data in certain data formats such as KML or KMZ (which is an compressed version of KML).

Austria provides the districts borders of Vienna at their [data.gv.at](https://www.data.gv.at/) page:

- https://www.data.gv.at/katalog/dataset/2ee6b8bf-6292-413c-bb8b-bd22dbb2ad4b

The ```.kml``` file can be retrieved under following url:

```{r}
districts_kml_url <- "https://data.wien.gv.at/daten/geo?version=1.3.0&service=WMS&request=GetMap&crs=EPSG:4326&bbox=48.10,16.16,48.34,16.59&width=1&height=1&layers=ogdwien:BEZIRKSGRENZEOGD&styles=&format=application/vnd.google-earth.kml+xml"

p_load(devtools)
if (!require("sf")) devtools::install_github("r-spatial/sf")

# Both the URL and the local path to the .kml-file work
vienna_districts <- read_sf(districts_kml_url)

vienna_districts$geometry
vienna_districts$geometry[1]
vienna_districts$Description

glimpse(vienna_districts)
```

### Feature extraction

The ```geometry``` contains both a single point to the district and a polygon shape.

```{r}
points <- st_collection_extract(vienna_districts, type = "POINT")
points %>% select(geometry)
```

```{r}
polygons <- st_collection_extract(vienna_districts, type = "POLYGON")
polygons %>% select(geometry)
```

### Casting 

The polygons can be converted to the desired type.

```{r}
st_cast(polygons, "LINESTRING")$geometry
```
### Simple plotting

The ```ggplot``` package, provides the function ```geom_sf``` to plot the polygon shapes.

```{r districts-simple}
ggplot(vienna_districts, aes(fill = Description)) + 
  geom_sf() + 
  theme(legend.position = "none")
```

### HTML Parsing

Unfortunately the ```Description``` of the polygon shapes of the districts are in a HTML format, which needs to be parsed. The ```rvest``` package provides this functionality.

```{r}
p_load(rvest)

vienna_districts$Description[1]
h <- read_html(vienna_districts$Description[1])
h %>% html_nodes(xpath = "/html/body/ul/li[2]/strong/span") %>% html_text()
h %>% html_nodes(xpath = "/html/body/ul/li[2]/span") %>% html_text()
h %>% html_elements("li")
names <- h %>% html_elements(".atr-name") %>% html_text2()
values <- h %>% html_elements(".atr-value") %>% html_text2()

description <- tibble(names, values)
kable(description)
```

### Widen Tibble Data Frame

```{r}
p_load(tidyr)

kable(pivot_wider(description, names_from = names, values_from = values))
```


```{r}
extract_description <- function(vienna_district_description_html){
  district_html <- read_html(vienna_district_description_html)
  names <- district_html %>% html_elements(".atr-name") %>% html_text2()
  values <- district_html %>% html_elements(".atr-value") %>% html_text2()
  description_long <- tibble(names, values)
  description_wide <- pivot_wider(description_long, names_from = names, values_from = values)
  description_wide
}
```

### Element-Wise Function Application

Using ```map_dfr``` of the ```purrr``` package, the newly created ```extract_description``` function will be applied to each observation of the ```Description``` feature in the ```vienna_districts``` tibble.

```{r}
p_load(purrr)
district_descriptions <- vienna_districts$Description %>% map_dfr(extract_description)
kable(district_descriptions)
```

```st_coordinates``` converts single points into (x, y)-coordinate pairs.

```{r}
points <- vienna_districts %>%
  st_collection_extract(type = "POINT") %>% 
  select(geometry) %>% 
  sf::st_coordinates(points)

geometries <- vienna_districts %>% 
  select(geometry)

districts <- district_descriptions %>% 
  bind_cols(geometries) %>%
  bind_cols(points)

kable(districts)
```

### Plotting 

```{r districts-with-legend, fig.width = 10}
ggplot(districts, aes(geometry = geometry, fill = NAMEK_NUM)) + 
  geom_sf() +
  ggtitle("Districts of Vienna")
```

```{r districts-labeled}
ggplot(districts) + 
  geom_sf(aes(geometry = geometry, fill = NAMEK_NUM)) +
  geom_text(aes(x = X, y = Y, label = BEZ), color = "white", size = 3) +
  ggtitle("Districts of Vienna") + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(legend.position = "none")
```


## References 

- https://ggplot2.tidyverse.org/reference/map_data.html
- https://github.com/edzer/sp/
- https://github.com/r-spatial/sf
- https://stackoverflow.com/questions/44320008/parse-html-data-using-r
- https://rvest.tidyverse.org/
- https://yutani.rbind.io/post/geom-sf-text-and-geom-sf-label-are-coming/
- https://www.youtube.com/watch?v=AgWgPSZ7Gp0
- https://www.youtube.com/watch?v=M7khlIRlr4s
- https://www.nceas.ucsb.edu/sites/default/files/2020-04/ggmapCheatsheet.pdf
