---
title: "You complete me"
output: rmarkdown::github_document
---

https://www.youtube.com/watch?v=ZLVWUGMkJCY

```{r}
if (!require("pacman")) install.packages("pacman")
library("pacman")

p_load(dplyr)
p_load(tidyr)
p_load(ggplot2)
p_load(hms)

knitr::opts_chunk$set(
  fig.path = "figures/you-complete-me/"
)
```

## 1 - Introduction

Lets say we have a data frame with prices for a specific product with different groups and the price for a given time:

```{r}
df1 <- tibble(
  product = c("Super", "Diesel", "Super", "Super", "Diesel"),
  time = c("00:00", "00:00", "01:00", "02:00", "02:00"),
  price = c(1.10, 2.10, 1.11, 1.12, 2.13)
)

p_load(knitr)
kable(df1)
```

*Note*: The price for Diesel at 01:00 AM was not recorded/ is missing.

The missing row for the Diesel price at 01:00 AM can be generated using the ```complete``` function:

```{r}
df1 %>% complete(product, time) %>% kable()
```

## 2 - Grouping

Let's assume the gas prices are associated with a unique gas station which is in a unique area. Area 1 always records the prices at the 00-minute mark of each hour of the day and area 2 always records the prices at the 01-minute mark of each of of the day.

```{r}
df2 <- tibble(
  area = c(1, 1, 2, 1, 2, 1, 1, 2),
  gas_station = c("A", "B", "C", "A", "C", "A", "B", "C"),
  time = c("00:00", "00:00", "00:01", "01:00", "01:01", "02:00", "02:00", "02:01"),
  price = c(1.10, 1.10, 1.05, 1.11, 1.04, 1.12, 1.12, 1.07)
)

df2 %>% kable()
```

*Note*: The price for gas_station B at 01:00 AM was not recorded/ is missing.

A simple ```complete``` would create an undesired result, because undesired rows would be generated. Since prices in area A are never recorded at the 01-minute mark, these rows are redundant:

```{r}
df2 %>% complete(area, gas_station, time) %>% kable()
```

To complete this date frame correctly, the price records have to be grouped by area.

```{r}
df2 %>% 
  group_by(area) %>% 
  complete(gas_station, time) %>%
  kable()
```

## 3 - Complex Grouping

Let's assume that each gas_station in each area records both the prices for Super and Diesel. The prices for each product of each area are reported at the same time.

```{r rows.print=20}
stations <- c("A", "B", "C")
products <- c("Super", "Diesel")
num_stations <- 3
num_products <- 2
num_records <- 3

set.seed(197)

df3 <- tibble(
  time = c(
    "00:00", "00:00", "00:01", "00:30", "00:30", "00:31", 
    "01:00", "01:00", "01:01", "01:30", "01:30", "01:31", 
    "02:00", "02:00", "02:01", "02:30", "02:30", "02:31"
    ), 
  area = rep(c(1, 1, 2), times = num_records * num_products),
  gas_station = rep(stations, times = num_products * num_records),
  product = rep(rep(products, each=num_stations), times = num_records),
  price = round(runif(n= num_records * num_products * num_stations, min=1, max=1.2), 2)
)

df3 %>% kable()
```

For whatever reason, some records have been lost:

```{r rows.print=20}
df3_missing <- df3 %>% 
  filter(!(time == "01:30" & gas_station == "B" & product == "Diesel")) %>%
  filter(!(time == "02:00" & gas_station == "A" & product == "Super")) %>%
  filter(!(time == "01:00" & gas_station == "A" & product == "Super"))

df3_missing %>% kable()
```

Now, to indicate the missing records, the missing rows can be generated using complete:

```{r rows.print=20}
df3_complete <- df3_missing %>% 
  group_by(area, product) %>% 
  complete(gas_station, time)
```

Now the tibble can be reordered again:

```{r rows.print=20}
df3_complete <- df3_complete %>% 
  relocate(time, area, gas_station, product) %>% 
  arrange(time, gas_station)

df3_complete %>% kable()
```

## 4 - Coupled columns

Let's say there are certain columns in the data frame that a coupled with each other
- product_name and product_id
- area_name and area_id
- station_name and station_id

```{r rows.print=20}
station_ids <- c(1, 2, 3)
station_names <- c("Shell", "BP", "ESSO")
area_ids <- c(1, 2)
area_names <- c("Berlin", "Frankfurt")
product_ids <- c(1, 2)
product_names <- c("Super", "Diesel")

num_stations <- length(station_ids)
num_products <- length(product_ids)
num_records <- 3

set.seed(345678)

df4 <- tibble(
  time = c(
    "00:00", "00:00", "00:01", "00:30", "00:30", "00:31", 
    "01:00", "01:00", "01:01", "01:30", "01:30", "01:31", 
    "02:00", "02:00", "02:01", "02:30", "02:30", "02:31"
    ), 
  area_id = rep(c(1, 1, 2), times = num_records * num_products),
  area = rep(c("Berlin", "Berlin", "Frankfurt"), times = num_records * num_products),
  
  station_id = rep(station_ids, times = num_products * num_records),
  station = rep(station_names, times = num_products * num_records),
  
  product_id = rep(rep(product_ids, each=num_stations), times = num_records),
  product = rep(rep(product_names, each=num_stations), times = num_records),
  
  price = round(runif(n= num_records * num_products * num_stations, min=1, max=1.2), 2)
)

df4 %>% kable()
```

And some records have been lost again:

```{r rows.print=20}
df4_missing <- df4 %>% 
  filter(!(time == "01:00" & station == "Shell" & product == "Super")) %>%
  filter(!(time == "01:30" & station == "BP" & product == "Diesel")) %>%
  filter(!(time == "02:00" & station == "Shell" & product == "Super"))
  
df4_missing %>% kable()
```

A simple complete would generate too many rows:

```{r rows.print=20}
df4_missing %>% 
  group_by(area, product) %>% 
  complete(station, station_id, area_id, product_id, time) %>%
  kable()
```

To find only the combinations (of station, station_id, area_id and product_id) that occur in the group data, ```nesting``` should be used:

```{r rows.print=20}
df4_complete <- df4_missing %>% 
  group_by(area, product) %>% 
  complete(nesting(station, station_id, area_id, product_id), time)

df4_complete %>% kable()
```

Now the tibble can be reordered again:

```{r rows.print=20}
df4_complete <- df4_complete %>% 
  relocate(time, area_id, area, station_id, station, product_id, product) %>% 
  arrange(time, station_id)

df4_complete %>% kable()
```
### Plot comparison

```{r original}
ggplot(df4, aes(parse_hm(time), price, color=station)) + 
  geom_point() +
  geom_line() + 
  xlab("time") +
  ggtitle("Orignal Data Frame") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r missing}
ggplot(df4_missing, aes(parse_hm(time), price, color=station)) + 
  geom_point() +
  geom_line() + 
  xlab("time") +
  ggtitle("Data Frame with missing Data Points") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r completed}
ggplot(df4_complete, aes(parse_hm(time), price, color=station)) + 
  geom_point() +
  geom_line() + 
  xlab("time") +
  ggtitle("Completed Data Frame with missing Data Points") +
  theme(plot.title = element_text(hjust = 0.5))
```


## References

- https://tidyr.tidyverse.org/reference/complete.html
- https://web.archive.org/web/20200312082724/http://www.imachordata.com/you-complete-me/

### Misc

- https://stackoverflow.com/questions/39814916/how-can-i-see-output-of-rmd-in-github/39816334
- http://uc-r.github.io/setting_seed/